### 2013-07-01 PLOS'13/evaluation.tex
updated a bit of evaluation text

To evaluate our system with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules. We implemented similar wrapping mechanism for DRK to attach and detach the framework at module interface. We checked the correctness of our system by running different filesystem and network modules(ext2/ext3, xfs, e1000e etc) under their control. To stress test the filesystem modules we used filesystem benchmark IOzone and ran the modules with null instrumentation client. null instrumentation client was used to avoid the overhead of instrumentation. We ran our tests on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ i7-860 2.80 GHz CPU, 8GB memory, and an Intel 82578DM Gigabit Ethernet card.