### 2013-09-29 thesis/abstract.tex

Finding, understanding, and fixing bugs in software systems is challenging. Dynamic binary translation (DBT) systems provide a powerful facility for building program analysis and debugging tools. However, DBT abstractions are too low-level and provide limited contextual information to instrumentation tools, making it hard to implement such tools.

In this theis, we introduce \emph{behavioral watchpoints}, a new software-based watchpoint framework that simplifies the implementation of DBT-based program analysis and debugging tools. Behavioral watchpoints have two key features: 1) they provide contextual information at the instruction level which are directly available with watchpoints and 2) they enable specializing instruction-level instrumentation with individual data structures. We describe three applications that were easily developed using our watchpoint framework: detecting buffer overflows, detecting read-before-write and memory freeing bugs and detecting memory leaks. We implemented behavioral watchpoints using Granary, a DBT framework for instrumenting operating system kernels. We evaluated the overheads of watchpoints for analyzing and debugging operating system kernel modules and show that these overheads are reasonable.