### 2013-11-11 EuroSys'14/abstract.tex

\documentclass[11pt,preprint,onecolumn,nocopyrightspace]{sigplanconf}
\usepackage{amsmath,epsfig,subfigure,multirow}

\usepackage{graphicx}
\usepackage{refstyle}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{framed}
\usepackage{lipsum}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage{numprint}
\usepackage[dvipsnames]{color}
\definecolor{shadecolor}{RGB}{250,250,250}

\begin{document}

\titlebanner{}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

\title{Trap or Trace: Implementing Kernel-space Data Watchpoints Efficiently}
\subtitle{}

\authorinfo{Peter Goodman\and Angela Demke Brown}
           {Department of Computer Science\\ University of Toronto}
           {\{pag, demke\}@cs.toronto.edu}

\authorinfo{Akshay Kumar\and Ashvin Goel}
           {Department of Electrical and Computer Engineering\\ University of Toronto}
           {\{akshay, ashvin\}@eecg.toronto.edu}

\maketitle

\begin{abstract}

Finding, understanding, and fixing bugs in operating system kernels is challenging. Watchpoints (also called data breakpoints) are a powerful tool that help with this problem; however, hardware-based watchpoints are too scarce for many uses and existing software-based watchpoints are either slow, or lack contextual information about the watched memory.

In this paper, we introduce \emph{behavioral watchpoints}: a new form of software-based watchpoints. Behavioral watchpoints extend on traditional watchpoints in novel ways: they watch ranges of addresses and they maintain context-specific information about watched memory, enabling debugging applications that cannot be easily implemented using traditional hardware or software watchpoints. For example, our framework can specialize instrumentation based on the context in which a watchpoint is triggered, which makes it possible to limit the scope of heavy-weight instrumentation to only code operating on watched data.

We discuss and evaluate several implementation strategies for our behavioral watchpoints framework, ranging from trap-based watchpoints (data-centric) to whole-kernel instrumentation based watchpoints (code-centric). We show that behavioral watchpoints, as implemented within a kernel dynamic binary translation framework, have reasonable overheads for their intended use in analysing and debugging operating system kernels.
\end{abstract}
\end{document}

%need to explain how the novelty helps (which is what I have tried above but the text is not strong).

%also, we complained earlier that software watchpoints are slow but we don't say that behavioural watchpoints are fast, so there is a disconnect there.

%also, the lack of contextual information, versus having contextual information issue, is hard to understand. Perhaps, saying that behavioral watchpoints enable specializing instrumentation based on context (and giving a simple example about light versus heavy weight instrumentation) might be easier for this audience to understand.